Redundancy analysis for hypertension

### 2019-05-15 partialrda.Rmd
Support for non-partial models

```{r data}
pseq.genus <- readRDS("data/phfinrisk_genus_all_drop50k_2018-11-16.RDs")
pseq.species <- readRDS("data/phfinrisk_species_all_drop50k_2018-12-21.RDs")
```

Updating phyloseq objects

```{r updating phyloseq objects}
phyloseq::sample_data(pseq.genus) <- phyloseq::sample_data(filter.phenotype.data(pseq.genus))
phyloseq::sample_data(pseq.species) <- phyloseq::sample_data(filter.phenotype.data(pseq.species))
pseq.genus.core <- core(pseq.genus, detection = 0.001, prevalence = 0.01)
```

```{r transformations}
rda.abundances.species  <- microbiome::transform(pseq.species, "clr") %>%
    microbiome::abundances() %>%
    t() %>%
    as.matrix()

rda.abundances.genus  <- microbiome::transform(pseq.genus, "clr") %>%
    microbiome::abundances() %>%
    t() %>%
    as.matrix()

rda.abundances.genus.core  <- microbiome::transform(pseq.genus.core, "clr") %>%
    microbiome::abundances() %>%
    t() %>%
    as.matrix()
```
## All models

```{r model function}
modelfunction <- function(x, rda.abund, df.meta) {
    covariates <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X",
                   "DIAB", "BL_USE_RX_C03", "BL_USE_RX_C07",
                   "BL_USE_RX_C08", "BL_USE_RX_C09")
    fo <- sprintf("rda.abund ~ %s + Condition(%s)", x, paste0(covariates, collapse = "+"))
    model <- rda(as.formula(fo), data = df.meta)
    anova.term <- anova(model, by="term", perm=1000)
    data.frame(var = as.character(x),
               beta = coef(model)[,1][[ifelse(x == "HYPERTENSION", "HYPERTENSION1", x)]],
               RDA.Conditioned = summary(model)$partial.chi,
               RDA.Constrained = summary(model)$constr.chi,
               RDA.Unconstrained = summary(model)$unconst.chi,
               p.anova = anova.term[["Pr(>F)"]][1],
               fo = fo)
}
```

```{r return parser}
returnparser <- function(ret){
    res <- do.call(rbind, ret) %>%
        mutate(RDA.total =  RDA.Conditioned + RDA.Constrained + RDA.Unconstrained,
               RDA.Conditioned.percent = RDA.Conditioned/RDA.total,
               RDA.Constrained.percent = RDA.Constrained/RDA.total,
               RDA.Unconstrained.percent = RDA.Unconstrained/RDA.total
               )

    res %>%
        dplyr::select(var,
                      beta,
                      p.anova,
                      RDA.Conditioned.percent,
                      RDA.Constrained.percent,
                      RDA.Unconstrained.percent) %>%
        kable()
}
```

@@

>>>>

```{r data}
pseq.genus <- import_filter_data("data/phfinrisk_genus_all_drop50k_2018-11-16.RDs")
pseq.species <- import_filter_data("data/phfinrisk_species_all_drop50k_2018-12-21.RDs")
pseq.genus.core <- core(pseq.genus, detection = 0.001, prevalence = 0.01)
```

```{r abundances to matrix function}
transform.matrix <- function(pseq, transformation = "clr") {
    microbiome::transform(pseq, transformation) %>%
        microbiome::abundances() %>%
            t() %>%
            as.matrix()
 }
```

```{r transformations}
rda.abundances.species  <- transform.matrix(pseq.species)
rda.abundances.genus  <- transform.matrix(pseq.genus)
rda.abundances.genus.core  <- transform.matrix(pseq.genus.core)
```
## All models

```{r model function}
modelfunction <- function(x,
                          rda.abund,
                          df.meta,
                          modelstr = "rda.abund ~ %s + Condition(%s)",
                          permutations = 1000) {
    covariates <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X",
                   "DIAB", "BL_USE_RX_C03", "BL_USE_RX_C07",
                   "BL_USE_RX_C08", "BL_USE_RX_C09")
    fo <- sprintf(modelstr, x, paste0(covariates, collapse = "+"))
    model <- rda(as.formula(fo), data = df.meta)
    anova.term <- anova(model, by="term", perm=permutations)
    data.frame(var = as.character(x),
               beta = coef(model)[,1][[ifelse(x == "HYPERTENSION", "HYPERTENSION1", x)]],
               conditioned = ifelse("partial.chi" %in% names(summary(model)),
                                        summary(model)$partial.chi,
                                        0),
               constrained = summary(model)$constr.chi,
               unconstrained = summary(model)$unconst.chi,
                p.anova = anova.term[["Pr(>F)"]][1],
               fo = fo)
}
```

```{r return parser}
returnparser <- function(ret){
    res <- do.call(rbind, ret) %>%
        mutate(RDA.total =  conditioned + constrained + unconstrained,
               conditioned = conditioned/RDA.total,
               constrained = constrained/RDA.total,
               unconstrained = unconstrained/RDA.total
               )

    res %>%
        dplyr::select(var,
                      beta,
                      p.anova,
                      conditioned,
                      constrained,
                      unconstrained) %>%
        kable()
}
```

@@

## Non-partial RDA species

```{r species non condition covariatesmodels test, eval = FALSE}
rda.subset <- rda.abundances.species[1:100, 1:100]
meta.subset <- meta(pseq.species) %>%
    tibble::rownames_to_column("id") %>%
    dplyr::filter(id %in% rownames(rda.subset))

ret.species <- mclapply(c2l("SYSTM", "DIASM", "PULSEPRESSURE", "MAP", "HYPERTENSION"),
              modelfunction,
              rda.abund = rda.subset,
              df.meta = meta.subset,
              modelstr = "rda.abund ~ %s + %s",
              mc.cores = 5)

returnparser(ret.species)
```
